---
title: "No shows self project"
author: "Ritika Agrawal"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a self project to find out the reasons for high rate of Doctor's appointment no-show in Brazil across 2014-15.
The Data Analytics techniques used in the project are data cleaning and Exploratory Data Analysis.

#### Loading in required Libraries
```{r}
suppressMessages(library(ggplot2))
suppressMessages(library(plyr))
suppressMessages(library(dplyr))
suppressMessages(library(reshape2))
suppressMessages(library(Hmisc))
suppressMessages(library(GGally))
```

#### Reading in Data
```{r}
no_shows <- read.csv("C:/Users/Dell/Documents/EDA/noshows.csv")
str(no_shows)
```

#### Convert No = 1 and Yes = 0
```{r}
no_shows['No.show'] <- lapply(no_shows['No.show'],factor,levels = c('Yes','No'),labels = c(0,1))
str(no_shows)
```

#### View initial few data values
```{r}
head(no_shows)
```

#### We can see that apponitment and scheduled days have timestamp and letter associated at the end, both of which are not relevant. Thus, we change these to only days.
```{r}
no_shows$AppointmentDay <- gsub(pattern = "[A-Z]",replacement = "",x = no_shows$AppointmentDay)
head(no_shows$AppointmentDay)
no_shows$AppointmentDay <- as.Date(no_shows$AppointmentDay)
head(no_shows$AppointmentDay)
no_shows$ScheduledDay <- gsub(pattern = "[A-Z]",replacement = "",x = no_shows$ScheduledDay)
no_shows$ScheduledDay <- as.Date(no_shows$ScheduledDay)
head(no_shows)
```

#### This looks good now! For further analysis, we might require Months separately, thus I'll extract year, month and date from appointment day.
```{r}
no_shows <- transform(no_shows,year=format(AppointmentDay,"%y"))
no_shows <- transform(no_shows,month=format(AppointmentDay,"%m"))
no_shows <- transform(no_shows,date=format(AppointmentDay,"%d"))
no_shows$year <- as.integer(as.character(no_shows$year))
no_shows$month <- as.integer(as.character(no_shows$month))
no_shows$date <- as.integer(as.character(no_shows$date))
head(no_shows)
```

#### I'll make the clumn names consistent and easy to refer using plyr.
```{r}
no_shows <- plyr::rename(no_shows,c("PatientId"="pat_id", "AppointmentID"="appt_id", "Gender"="gender", "ScheduledDay"="sch_day", "AppointmentDay"="appt_day", "Age"="age", "Neighbourhood"="neighbour", "Scholarship"="scholarship", "Hipertension"="hypertension", "Diabetes"="diabetes", "Alcoholism"="alcoholism", "Handcap"="handicap", "SMS_received"="sms", "No.show"="noshow", "year"="appt_year", "month"="appt_month", "date"="appt_date"))
no_shows$noshow <- as.integer(as.character(no_shows$noshow))
head(no_shows)
```

#### Let's look at the summary statistics of our data.
```{r}
summary(no_shows)
```

#### From summary, we observe that age gives minimum value as -1! Thus, here comes DATA CLEANING. I'll remove the rows from data set which have age<0. Note that we keep the age = 0 to account for the fact that there might be new born kids who are less than one year of age.
```{r}
no_shows_updatedage <- subset(no_shows, age>=0)
summary(no_shows_updatedage)
```
#### Thus, we see, minimum age now becomes 0.
#### Since there is no more anomalies with dataset, like negative values or NaN, we can move to EDA.

### Exploratory Data Analysis


